Dish: Distribution-oblivious Shell Scripting

### 2020-03-03 pldi/draft.tex
A pass on 5.1

  \item
    \S\ref{impl} describes the \sys implementation and addresses several challenges related to the translation and orchestration of distributed shell programs.
    %% details other concerns, such as distributed state management and extensibility. 
\end{itemize}

@@

This section describes challenges related to the implementation of \sys. 
It starts by describing how the \sys front-end translates a shell script to the dataflow graph model defined in \S\ref{ir}, by applying transformation on distributable regions~\sx{front-end}.
It closes with the \sys planner, which  optimizes, distributes, and orchestrates the execution of a given dataflow graph~\sx{optimizer}.

@@

As mentioned in \Cref{ir}, in order to generate the distributed implementation of a shell script, \sys translates it to a dataflow graph.
However, there are several components in shell programs that cannot be arbitrarily parallelized without affecting the program behaviour---\eg commands connected with a \ttt{\&\&} operator~\sx{bg}.
The fact that the shell is highly dynamic compounds the challenge---\eg by not knowing what files does an unexpanded command argument refer to.
% of correctly identifying what parts can be distributed, e.g.

@@

To address this issue, we introduce the notion of distributable program regions and a translation pass for converting them to a dataflow graph.
The translation pass identifies distributable regions and translates each of them to a dataflow graph, provided the abstract syntax tree (AST) of a shell script.
The AST is produced by \ttt{LibDash}, a POSIX compliant parser~\cite{libdash}.

\paragraph{Distributable Regions}
%
%% \kk{The goal of this subsection is to explain that Dish does a safe
%%   but effective analysis to identify which parts of the scripts to
%%   parallelize.}
%
Distributable regions are (ideally, maximal) program sub-expressions that can be parallelized safely with respect to their sequential execution.
The structure of a shell program can be used to identify components that can be executed concurrently and barriers that are natural points for enforcing synchronization.
Consider the following example:

@@

\noindent
The commands \ttt{cat f1 f2} and \ttt{grep ``foo''} would already execute concurrently in the standard shell, while \ttt{sort f3} would wait for their completion before executing.
Both \ttt{cat f1 f2 | grep ``foo'' > f3} and \ttt{sort f3} are distributable regions---maximal sub-expressions that cannot be extended beyond \ttt{\&\&}.

Intuitively, distributed regions correspond to subsections of the program that would be allowed to execute concurrently by different processes in the POSIX standard~\cite{posix}.
The pipe operator \ttt{|} and the parallel composition operator \ttt{\&} enable parallelization, and extend distributable regions.
Conversely, the \ttt{;} sequential composition operator, the \ttt{\&\&} logical operator, and function definitions represent barrier constructs that do not allow distributable regions to permeate through.

@@

\paragraph{Translation Pass}
The \sys front-end performs a depth-first search pass on the AST of the given shell program.
During this pass, it greedily extends the distributable regions from the bottom up, translating their independent components to dataflow graph nodes until a barrier construct is reached.
To identify possible opportunities for parallelism, the translation pass also extracts the category of each node together with its input and output pairs.
To perform this extraction, it uses the command name to index into a key-value map with values that are expressed in the DSL for distributability characterization~\sx{ext}.

>>>>

  \item
    \S\ref{impl} describes the \sys implementation and addresses
    several challenges related to the translation and optimization of
    shell programs.
    %% details other concerns, such as distributed state management and extensibility. 
\end{itemize}

@@

This section describes challenges related to the implementation of
\sys. It starts by describing how the \sys front-end translates a
shell script to the dataflow graph model defined in \S\ref{ir}, by
applying transformation on parallelizable regions~\sx{front-end}.  It
closes with the \sys optimizer, which optimizes the dataflow graph
exposing data-parallelism, and outputs an optimized
parallel script~\sx{optimizer}.

@@

As mentioned in \Cref{ir}, in order to optimize a shell script, \sys
translates it to a dataflow graph.  However, there are several
components in shell programs that cannot be arbitrarily parallelized
without affecting the program behaviour---\eg commands connected with
a \ttt{\&\&} operator~\sx{bg}. In addition, the shell's highly dynamic
nature compounds the challenge---\eg by not knowing what files does an
unexpanded command argument refer to.
% of correctly identifying what parts can be distributed, e.g.

@@

To address the first issue, we introduce the notion of parallelizable
program regions and a translation pass for converting them to a
dataflow graph. To address the second issue, we limit dynamic
components by deferring translation as late as possible using a
wrapping scheme.

%% The translation pass identifies distributable regions and
%% translates each of them to a dataflow graph, provided the abstract
%% syntax tree (AST) of a shell script.  The AST is produced by
%% \ttt{LibDash}, a POSIX compliant parser~\cite{libdash}.

\paragraph{Parallelizable Regions}
%
%% \kk{The goal of this subsection is to explain that Dish does a safe
%%   but effective analysis to identify which parts of the scripts to
%%   parallelize.}
%
Parallelizable regions are program sub-expressions that can be
parallelized safely without altering the total program behaviour.
%% respect to their sequential execution.
The search for these regions can be guided by the structure of shells
programs, which inherently contains information about components that
can be executed independently and barriers that are natural points for
enforcing synchronization. Consider the following example:

@@

\noindent
The commands \ttt{cat f1 f2} and \ttt{grep ``foo''} would execute as
independent processes in the standard shell, while \ttt{sort f3} would
wait for their completion before executing.  Both \ttt{cat f1 f2 |
  grep ``foo'' > f3} and \ttt{sort f3} are therefore parallelizable
regions, regions that cannot be extended beyond \ttt{\&\&}---even
better, they are maximal.

Intuitively, parallelizable regions correspond to sub-expressions of
the program that would be allowed to execute independently by
different processes in the POSIX standard~\cite{posix}. Larger
parallelizable regions can be composed from smaller ones using the
pipe operator \ttt{|} and the parallel composition operator \ttt{\&}.
Conversely, the \ttt{;} sequential composition operator, the
\ttt{\&\&} logical operator, and function definitions represent
barrier constructs that do not allow parallelizable regions to
permeate through.

@@

\paragraph{Translation Pass}

\kk{It might be a good idea to change this subsection based on the new
  as-late-as-possible architecture where everything is wrapped and
  translated at the final moment. Ideally, this would become the main
  point of this subsection, as the translation algorithm is kind of
  redundant when having the distributable regions too.}

The \sys front-end performs a depth-first search pass on the AST of
the given shell program.  During this pass, it greedily extends the
distributable regions from the bottom up, translating their
independent components to dataflow graph nodes until a barrier
construct is reached.  To identify possible opportunities for
parallelism, the translation pass also extracts the category of each
node together with its input and output pairs.  To perform this
extraction, it uses the command name to index into a key-value map
with values that are expressed in the DSL for distributability
characterization~\sx{ext}.
